Solar wind predictions for the Parker Solar Probe orbit

### 2017-10-25 paperMVVB/sections/content.tex
#4 integrated changes by MVVB; improved figures

%transition
For the bulk of the solar wind these static lognormal functions describe the parameters' distributions well\textbf{. The abnormally high parameter values in the distribution functions can be attributed to shock/CME events in agreement with the results of the OMNI solar wind investigations by \citet{Richardson2012}.} The simple lognormal fit functions underestimate the \textbf{frequencies in} their high value tails, except for the temperature’s tail which is overestimated, as seen in the insets of Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}\textbf{, this appears to be because CMEs do not come with abnormally high temperatures.} The velocity's compositional lognormal fit only slightly overestimates its tail as seen in the inset of Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}.
The slow and fast part contribute almost equally ($c \approx 0.5$) to the long-term velocity distribution function.

@@

\textbf{The radial dependencies of the key solar wind parameters over the distance range \SIrange{0.29}{0.98}{\au} measured by both Helios probes are plotted in Fig.~\ref{fig:radial_fit_4_thesis_light_b_skip}, together with their median and mean values for different solar distances, calculated for the minimal distance resolution \SI{0.01}{\au} of the data set.}
\begin{figure*}
	\includegraphics[width=18cm]{figures/radial_fit_4_thesis_light_b_skip.pdf}
	\caption{Helios hourly data plots of the four solar wind parameters over solar distance. The mean and median per \SI{0.01}{\au} data bin and their fit curves are plotted as well. The Helios data has a native distance resolution of \SI{0.01}{\au}, thus, to make the abundance visible in these plots, we added a random distance value of up to \SI{+-0.005}{\au}. The higher velocity data points above \SI{800}{\km\per\s} \hl{identified as CME events by BS1998, Sheeley1981 are circled red.}}
	\label{fig:radial_fit_4_thesis_light_b_skip}
\end{figure*}
Assuming a radial solar wind outflow, it is expected that the distance dependence of the solar wind parameters over the Helios data range \SIrange{0.29}{0.98}{\au} can be described through power law scaling. Therefore we use the power law function
\begin{align}
	x(r) = d\cdot r^e	\label{eq:power_function}
\end{align}
for the regression fit of the median and mean, with $r$ being the solar distance in astronomical units, $d$ the magnitude at \SI{1}{\au} and $e$ the exponent. The fits are weighted through the different data counts per bin.
%fit result table and figure
\textbf{The obtained coefficients for the median and mean power law fits ($d_\text{med}$, $e_\text{med}$, $d_\text{avg}$ and $e_\text{avg}$) are listed in Table~\ref{tab:mean_median_fit_parameter} and their corresponding curves are shown in Fig.~\ref{fig:radial_fit_4_thesis_light_b_skip}.}
\begin{table*}
	\caption{Fit coefficients for the median and mean solar distance dependencies \textbf{(\ref{eq:power_function})} of the four solar wind parameters derived from the combined Helios~1 and 2 data. \textbf{The numbers in parentheses are the estimated standard deviations of the fit parameters referred to the corresponding last digits of the quoted value.} The crossing distances indicate where the median and mean fits intersect each other. The yearly variation is the weighted standard deviation derived from the yearly fit exponents seen in Fig.~\ref{fig:yearly_gradients_c}.}
	\label{tab:mean_median_fit_parameter}
	\centering
	\begin{tabular}{l
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	@{}c
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	S[table-format = 1.1(2)e1]
	S[table-format = 1.3]}
		\hline\hline
		\multirow{2}{*}{Parameter}	&\multicolumn{2}{c}{Median}	&	&\multicolumn{2}{c}{Mean}	&\multicolumn{1}{c}{Crossing distance}	&\multicolumn{1}{c}{Yearly variation}\\
		\cline{2-3}	\cline{5-6}
			&\multicolumn{1}{c}{$d_\text{med}$}	&\multicolumn{1}{c}{$e_\text{med}$}	&	&\multicolumn{1}{c}{$d_\text{avg}$}	&\multicolumn{1}{c}{$e_\text{avg}$}	&\multicolumn{1}{c}{[\si{\au}]}	&\multicolumn{1}{c}{$\Delta e$}\\
		\hline
		\textbf{Magnetic field [\si{nT}]}	&5.377(92)	&-1.655(17)	&	&6.05(10)	&-1.546(18)	&0.339(11)	&0.11\\
		\textbf{Velocity [\SI{e2}{\km\per\s}]}	&4.107(28)	&0.058(13)	&	&4.356(24)	&0.049(10)	&0.7(83)e3	&0.012\\
		\textbf{Density [\si{\per\cm\cubed}]}	&5.61(27)	&-2.093(46)	&	&7.57(30)	&-2.010(38)	&0.027(73)	&0.072\\
		\textbf{Temperature [\SI{e4}{\K}]}	&7.14(23)	&-0.913(39)	&	&9.67(21)	&-0.792(28)	&0.082(85)	&0.050\\
		\hline
	\end{tabular}
\end{table*}
% source year variation: fit_out_yearly_gradients_table_b_stats.txt

@@

%crossing distances
\textbf{The mean and median velocity fit exponents acquired from the Helios data} are very similar, which indicates that they just as well can be kept identical so that the basic shape of the frequency distribution does not change with distance. Contrary, the mean and median fits for the magnetic field strength cross each other at \SI{0.339}{\au} \textbf{(see Table~\ref{tab:mean_median_fit_parameter})} and the mean \hl{is \textbf{getting slightly} lower than the median at smaller distances }. Thus, below that distance the \textbf{frequency distribution} cannot well be described anymore by a lognormal function\textbf{, because the mean of a lognormal function has to be larger than its median (as mentioned in Sect.~\ref{sec:frequency_distribution}), that is, the location of the crossing indicates that the parameter's distribution is not anymore of a lognormal shape thereafter.} The fits for the proton temperature show a similar behavior, having an \textbf{extrapolated} intersection at \SI{0.082}{\au}. Therefore the extrapolation of the magnetic field and temperature distribution frequencies to the PSP orbit by applying lognormal functions is limited. \textbf{The crossing points limit the regions where the distribution's shapes can still be considered lognormal.}

@@

%\subsection{Near-Sun extrapolation for PSP orbit}
%PSP orbit
Parker Solar Probe is planned to launch in mid 2018. With its first Venus flyby it will swing into Venus' orbital plane, reaching already 93~days after launch in November 2018 a first perihelion with a distance of \SI{0.16}{\au}. Seven additional Venus flybys allow \textbf{the perihelion distance to be reduced} to a minimum of \SI{9.86}{\Rs}\textbf{. This distance will be reached with the 22nd perihelion in December 2024 \citep{Fox2015} as plotted in the top panel of Fig.~\ref{fig:SPP_orbit_predicted_SSN_overview_f_plot}}.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_orbit_predicted_SSN_overview_f_plot.pdf}}
	\caption{PSP's solar distance during its mission time (top). Consecutive Venus flybys bring its perihelia nearer to the Sun. Actual and predicted SSN (bottom), that is, SIDC 13-month smoothed monthly actual SSN, SIDC prediction, SWPC prediction and by 11~years shifted SSN from previous cycle~24, together with two alternative trends of half and twice its amplitude.}
	\label{fig:SPP_orbit_predicted_SSN_overview_f_plot}
\end{figure}

We extrapolate the derived empirical solar wind \textbf{models} to PSP’s orbital distance range and compare the results with those from the existing models shown in Fig.~\ref{fig:sw_extrapolation_ssn_e_plot}.

@@

	\includegraphics[width=18cm]{figures/sw_extrapolation_ssn_e_plot.pdf}

@@

	\label{fig:sw_extrapolation_ssn_e_plot}

@@

%\subsection{SSN prediction for PSP mission time}
%SSN prediction
\textbf{Aside from the solar distance the derived solar wind parameter models depend on the SSN. Short-term predictions of the SSN can be used for the solar wind predictions of PSP's early perihelia and also for refining the solar wind predictions during PSP's mission.} Several sources are available for SSN short-term predictions. The SIDC provides 12-month SSN forecasts\footnote{\url{http://sidc.be/silso/forecasts}} obtained from different methods (e.g., Kalman filter combined method). The SSN prediction of NOAA's Space Weather Prediction Center (SWPC) follows for the time period until end of 2019 a consensus of the Solar~Cycle~24 Prediction~Panel\footnote{\url{http://www.swpc.noaa.gov/products/solar-cycle-progression}}.
The SSN for PSP's first perihelion will be small---certainly below 20, whereas \textbf{the SSN during the closest perihelia, which will commence at the end of 2024 at the likely maximum phase of cycle~25, cannot be predicted at this time.} \textbf{However \citet{Hathaway2016} found indications that the next solar cycle will be similar in size to the current cycle~24. Therefore we simply assume a pattern similar to the last cycle for the prediction of the next solar} cycle and thus shift the last cycle by 11~years. Additionally we consider as possible alternatives SSN patterns of half and twice its amplitude as shown in \textbf{the bottom panel of Fig.~\ref{fig:SPP_orbit_predicted_SSN_overview_f_plot}.}

%\subsection{PSP solar wind environment estimation}
Implementing the \textbf{SSN predictions} for the PSP mission time and the orbital trajectory data, \textbf{we can infer} which solar wind parameter magnitudes can be expected. Figs.~\ref{fig:SPP_perihelia_prediction_f_plot} and \ref{fig:SPP_perihelia_prediction_nearest_f_plot} show \textbf{the median values (\ref{eq:median_with_ssn}) of} the considered different solar wind parameters for 12-day periods, comprising the first perihelion in Novemver 2018 and the \textbf{first} closest perihelion in December 2024.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_perihelia_prediction_f_plot.pdf}}
	\caption{Estimated solar wind parameter medians (black lines) and their error bands (grey) during 12~days in 2018 with PSP's first perihelion at about \SI{0.16}{\au}. For the velocity the combined median is calculated and also the SSN independent slow and fast parts are plotted (dotted lines).}
	\label{fig:SPP_perihelia_prediction_f_plot}
\end{figure}
% SSN sources and error ranges for first perihelion
% magnetic field:
% SWPC prediction - 0 years
% SWPC expected SSN ranges
% velocity:
% actual monthly SSN - 3 years
% not plotted
% density:
% actual monthly SSN - 6 years
% SIDC given std. dev. from stations
% temperature:
% actual monthly SSN - 3 years
% SIDC given std. dev. from stations
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_perihelia_prediction_nearest_f_plot.pdf}}
	\caption{Estimated solar wind parameter medians (black lines) and their error bands (grey) \textbf{during} 12~days in 2024 with PSP's \textbf{22nd (the first closest)} perihelion at \SI{0.0459}{\au}. For the velocity the combined median is calculated and also the SSN independent slow and fast parts are plotted (dotted lines).}
	\label{fig:SPP_perihelia_prediction_nearest_f_plot}
\end{figure}
% SSN sources and error ranges for first closest perihelion
% magnetic field:
% actual monthly SSN - 11 years
% half/twice SSN
% velocity:
% actual monthly SSN - 14 years
% not plotted
% density:
% SWPC prediction - 0 years
% SWPC expected SSN ranges
% temperature:
% actual monthly SSN - 14 years
% half/twice SSN
In the beginning of the mission median values of about \SI{87}{\nT}, \SI{340}{\km\per\s}, \SI{214}{\per\cm\cubed} and \SI{503000}{\K} are estimated to be measured at \SI{0.16}{\au}, increasing to about \SI{943}{\nT}, \SI{290}{\km\per\s}, \SI{2951}{\per\cm\cubed} and \SI{1930000}{\K} during the \textbf{first} closest approach at \SI{0.046}{\au}. \textbf{Monthly SSNs---shifted by the time lags specific to the solar wind parameters---are used in the calculation of the solar wind predictions. These SSNs are either actual smoothed values from the SIDC with their reported standard deviations, short-term predictions from the SWPC with their expected ranges or actual smoothed values from the SIDC shifted by 11~years with half/twice their values as uncertainties. The error bands given in both figures, calculated from error propagation, include these SSN ranges and the derived fit parameter errors.}
%B
% 87.3642361846083
%N
% 213.613795298952	first
% 2950.67837375615	nearest
% 2925.63248639046	SSN 0
% 3421.59065204775	SSN 200
%T
% 503136.759398494

\textbf{Finally the estimated solar wind environment can be derived from the function $W'''(x,ssn,r)$. The estimated frequency distributions of the four solar wind parameters at PSP's first and 22nd (first closest) perihelia are plotted in Fig.~\ref{fig:SPP_sw_distributions}.}
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_sw_distributions.pdf}}
	\caption{\textbf{Frequency distributions of the four solar wind parameters (same as in Figs.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot} and \ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}) and those estimated with the solar wind model for PSP's first and 22nd (first closest) perihelion. In these figures the frequencies of both extrapolated curves are scaled for visibility to the same height as the \SI{1}{\au} distribution.}}
	\label{fig:SPP_sw_distributions}
\end{figure}
\textbf{Note again that the velocity and temperature distributions for the 22nd perihelion are only upper limits and the actual by PSP to be encountered values are expected to be smaller.}

@@

\textbf{For determining solar activity and solar distance dependent relations for the median and average solar wind values, we could have used the simpler approach of combining the radial dependence of averaged Helios data with averaged \SI{1}{\au} OMNI data scaled with the SSN. It is expected that the results of a simpler analysis would have similar distance scaling results, as can be inferred from the exponents in Tables~\ref{tab:mean_median_fit_parameter} and \ref{tab:extrapolation_model_fit_parameters}. However, in our study we are not only interested in averages but rather in bulk distributions, that is, the whole range of values that might occur. For the determination of the frequency distributions the use of the more complex fit model is important, because the distance between median and average values determines the width of the lognormal distributions. With our approach we are able to control and fit this distance.}

\textbf{It} is clear that the calculated distribution functions just represent first order estimates of the real solar wind to be encountered by PSP. The solar wind environment to be encountered will depend at times of PSP on the structure of the solar corona and underlying photospheric magnetic field and on the evolution and interaction of individual solar wind streams and superimposed CMEs and shocks. However, the derived results are in good agreement with existing studies about near-Sun solar wind magnetic field strenghts and densities as shown in Sect.~\ref{sec:model_extrapolation_to_psp_orbit}. \textbf{The extrapolation results of the velocity and the temperature differ from the direct measurements seen in existing studies. This suggests that below about \SI{20}{\Rs} PSP may dive into the region where the acceleration and heating of the solar wind is expected to occur} (see Fig.~\ref{fig:sw_extrapolation_ssn_e_plot}). The near-Sun solar wind velocity at PSP perihelion is also expected to be slower than our model estimates, because the \textbf{solar wind is assumed to be accelerated up to the height of the Alfvénic critical surface, which is predicted to lie on average around \SI{17}{\Rs} \citep[e.g.,][]{Sittler1999,Exarhos2000}, scaling with solar activity within a range of between \SI{15}{\Rs} at solar minimum and \SI{30}{\Rs} at solar} maximum \citep{Katsikas2010,Goelzer2014}.	% in average around 17~Rs (e.g., 0.6~Rs Schatten 1969, 17~Rs Sittler1999, 17~Rs Exarhos2000) and scales with solar activity between 15~Rs in minimum and 30~Rs in maximum (19~Rs Katsikas2010, 15–30~Rs Goelzer2014).

>>>>

%transition
For the bulk of the solar wind these static lognormal functions describe the parameters' distributions well\textbf{. The abnormally high parameter values in the distribution functions can be attributed to shock/CME events in agreement with the results of the OMNI solar wind investigations by \citet{Richardson2012}.} The simple lognormal fit functions underestimate the \textbf{frequencies in} their high value tails, except for the temperature’s tail which is overestimated, as seen in the insets of Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}\textbf{, this appears to be because CMEs do not come with abnormally high temperatures, but rather with temperatures lower than those of the average solar wind \citep{Forsyth2006}.} The velocity's compositional lognormal fit only slightly overestimates its tail as seen in the inset of Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}.
The slow and fast part contribute almost equally ($c \approx 0.5$) to the long-term velocity distribution function.

@@

\textbf{The radial dependencies of the key solar wind parameters over the distance range \SIrange{0.29}{0.98}{\au} measured by both Helios probes are plotted in Fig.~\ref{fig:radial_fit_4_thesis_light_b_skip}, together with their median and mean values for different solar distances, calculated for the minimal distance resolution \SI{0.01}{\au} of the data set.}
\begin{figure*}
	\includegraphics[width=18cm]{figures/radial_fit_4_thesis_light_b_skip.pdf}
	\caption{Helios hourly data plots of the four solar wind parameters over solar distance. The mean and median per \SI{0.01}{\au} data bin and their fit curves are plotted as well. The Helios data has a native distance resolution of \SI{0.01}{\au}, thus, to make the abundance visible in these plots, we added a random distance value of up to \SI{+-0.005}{\au}. \textbf{The high velocity data points above \SI{800}{\km\per\s} (circled red) are identified as CME events \citep[e.g.,][]{Bothmer1998}.}}
	\label{fig:radial_fit_4_thesis_light_b_skip}
\end{figure*}
Assuming a radial solar wind outflow, it is expected that the distance dependence of the solar wind parameters over the Helios data range \SIrange{0.29}{0.98}{\au} can be described through power law scaling. Therefore we use the power law function
\begin{align}
	x(r) = d\cdot r^e	\label{eq:power_function}
\end{align}
for the regression fit of the median and mean, with $r$ being the solar distance in astronomical units, $d$ the magnitude at \SI{1}{\au} and $e$ the exponent. The fits are weighted through the different data counts per bin.
%fit result table and figure
\textbf{The obtained coefficients for the median and mean power law fits ($d_\text{med}$, $e_\text{med}$, $d_\text{avg}$ and $e_\text{avg}$) are listed in Table~\ref{tab:mean_median_fit_parameter} and their corresponding curves are shown in Fig.~\ref{fig:radial_fit_4_thesis_light_b_skip}.}
\begin{table*}
	\caption{Fit coefficients for the median and mean solar distance dependencies \textbf{(\ref{eq:power_function})} of the four solar wind parameters derived from the combined Helios~1 and 2 data. \textbf{The numbers in parentheses are the estimated standard deviations of the fit parameters referred to the corresponding last digits of the quoted value.} The crossing distances indicate where the median and mean fits intersect each other. The yearly variation is the weighted standard deviation derived from the yearly fit exponents seen in Fig.~\ref{fig:yearly_gradients_c}.}
	\label{tab:mean_median_fit_parameter}
	\centering
	\begin{tabular}{l
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	@{}c
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	S[table-format = 1.1(2)e1]
	S[table-format = 1.3]}
		\hline\hline
		\multirow{2}{*}{Parameter}	&\multicolumn{2}{c}{Median}	&	&\multicolumn{2}{c}{Mean}	&\multicolumn{1}{c}{Crossing distance}	&\multicolumn{1}{c}{Yearly variation}\\
		\cline{2-3}	\cline{5-6}
			&\multicolumn{1}{c}{$d_\text{med}$}	&\multicolumn{1}{c}{$e_\text{med}$}	&	&\multicolumn{1}{c}{$d_\text{avg}$}	&\multicolumn{1}{c}{$e_\text{avg}$}	&\multicolumn{1}{c}{[\si{\au}]}	&\multicolumn{1}{c}{$\Delta e$}\\
		\hline
		\textbf{Magnetic field [\si{nT}]}	&5.377(92)	&-1.655(17)	&	&6.05(10)	&-1.546(18)	&0.339(11)	&0.11\\
		\textbf{Velocity [\SI{e2}{\km\per\s}]}	&4.107(28)	&0.058(13)	&	&4.356(24)	&0.049(10)	&0.7(83)e3	&0.012\\
		\textbf{Density [\si{\per\cm\cubed}]}	&5.61(27)	&-2.093(46)	&	&7.57(30)	&-2.010(38)	&0.027(73)	&0.072\\
		\textbf{Temperature [\SI{e4}{\K}]}	&7.14(23)	&-0.913(39)	&	&9.67(21)	&-0.792(28)	&0.082(85)	&0.050\\
		\hline
	\end{tabular}
\end{table*}
% source year variation: fit_out_yearly_gradients_table_b_stats.txt

@@

%crossing distances
\textbf{The mean and median velocity fit exponents acquired from the Helios data} are very similar, which indicates that they just as well can be kept identical so that the basic shape of the frequency distribution does not change with distance. Contrary, the mean and median fits for the magnetic field strength cross each other at \SI{0.339}{\au} \textbf{(see Table~\ref{tab:mean_median_fit_parameter})} and the mean is \textbf{getting slightly} lower than the median at smaller distances. Thus, below that distance the \textbf{frequency distribution} cannot well be described anymore by a lognormal function\textbf{, because the mean of a lognormal function has to be larger than its median (as mentioned in Sect.~\ref{sec:frequency_distribution}), that is, the location of the crossing indicates that the parameter's distribution is not anymore of a lognormal shape thereafter.} The fits for the proton temperature show a similar behavior, having an \textbf{extrapolated} intersection at \SI{0.082}{\au}. Therefore the extrapolation of the magnetic field and temperature distribution frequencies to the PSP orbit by applying lognormal functions is limited. \textbf{The crossing points limit the regions where the distribution's shapes can still be considered lognormal.}

@@

%\subsection{Near-Sun extrapolation for PSP orbit}
%PSP orbit
Parker Solar Probe is planned to launch in mid 2018. With its first Venus flyby it will swing into Venus' orbital plane, reaching already 93~days after launch in November 2018 a first perihelion with a distance of \SI{0.16}{\au}. Seven additional Venus flybys allow \textbf{the perihelion distance to be reduced} to a minimum of \SI{9.86}{\Rs}\textbf{. This distance will be reached with the 22nd perihelion in December 2024 \citep{Fox2015} as plotted in the top panel of Fig.~\ref{fig:SPP_orbit_predicted_SSN_overview_f_plot}}.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_orbit_predicted_SSN_overview_f_plot.pdf}}
	\caption{PSP's solar distance during its mission time (top). Consecutive Venus flybys bring its perihelia nearer to the Sun. Actual and predicted SSN (bottom), that is, SIDC 13-month smoothed monthly actual SSN, SIDC \textbf{Standard Curves Kalman filter} prediction, SWPC prediction and by 11~years shifted SSN from previous cycle~24, together with two alternative trends of half and twice its amplitude.}
	\label{fig:SPP_orbit_predicted_SSN_overview_f_plot}
\end{figure}

We extrapolate the derived empirical solar wind \textbf{models} to PSP’s orbital distance range and compare the results with those from the existing models shown in Fig.~\ref{fig:sw_extrapolation_ssn_f_plot}.

@@

	\includegraphics[width=18cm]{figures/sw_extrapolation_ssn_f_plot.pdf}

@@

	\label{fig:sw_extrapolation_ssn_f_plot}

@@

%\subsection{SSN prediction for PSP mission time}
%SSN prediction
\textbf{Aside from the solar distance the derived solar wind parameter models depend on the SSN. Short-term predictions of the SSN can be used for the solar wind predictions of PSP's early perihelia and also for refining the solar wind predictions during PSP's mission.} Several sources are available for SSN short-term predictions. The SIDC provides 12-month SSN forecasts\footnote{\url{http://sidc.be/silso/forecasts}} obtained from different methods (e.g., Kalman filter \textbf{Standard Curve} method). The SSN prediction of NOAA's Space Weather Prediction Center (SWPC) follows for the time period until end of 2019 a consensus of the Solar~Cycle~24 Prediction~Panel\footnote{\url{http://www.swpc.noaa.gov/products/solar-cycle-progression}}.
The SSN for PSP's first perihelion will be small---certainly below 20, whereas \textbf{the SSN during the closest perihelia, which will commence at the end of 2024 at the likely maximum phase of cycle~25, cannot be predicted at this time.} \textbf{However \citet{Hathaway2016} found indications that the next solar cycle will be similar in size to the current cycle~24. Therefore we simply assume a pattern similar to the last cycle for the prediction of the next solar} cycle and thus shift the last cycle by 11~years. Additionally we consider as possible alternatives SSN patterns of half and twice its amplitude as shown in \textbf{the bottom panel of Fig.~\ref{fig:SPP_orbit_predicted_SSN_overview_f_plot}.}

%\subsection{PSP solar wind environment estimation}
Implementing the \textbf{SSN predictions} for the PSP mission time and the orbital trajectory data, \textbf{we can infer} which solar wind parameter magnitudes can be expected. Figs.~\ref{fig:SPP_perihelia_prediction_f_plot} and \ref{fig:SPP_perihelia_prediction_nearest_f_plot} show \textbf{the median values (\ref{eq:median_with_ssn}) of} the considered different solar wind parameters for 12-day periods, comprising the first perihelion in Novemver 2018 and the \textbf{first} closest perihelion in December 2024.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_perihelia_prediction_f_plot.pdf}}
	\caption{Estimated solar wind parameter medians (black lines) and their error bands (grey) during 12~days in 2018 with PSP's first perihelion at about \SI{0.16}{\au}. For the velocity the combined median is calculated and also the SSN-independent slow and fast parts are plotted (dotted lines).}
	\label{fig:SPP_perihelia_prediction_f_plot}
\end{figure}
% SSN sources and error ranges for first perihelion
% magnetic field:
% SWPC prediction - 0 years
% SWPC expected SSN ranges
% velocity:
% actual monthly SSN - 3 years
% not plotted
% density:
% actual monthly SSN - 6 years
% SIDC given std. dev. from stations
% temperature:
% actual monthly SSN - 3 years
% SIDC given std. dev. from stations
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_perihelia_prediction_nearest_f_plot.pdf}}
	\caption{Estimated solar wind parameter medians (black lines) and their error bands (grey) \textbf{during} 12~days in 2024 with PSP's \textbf{22nd (the first closest)} perihelion at \SI{0.0459}{\au}. For the velocity the combined median is calculated and also the SSN-independent slow and fast parts are plotted (dotted lines).}
	\label{fig:SPP_perihelia_prediction_nearest_f_plot}
\end{figure}
% SSN sources and error ranges for first closest perihelion
% magnetic field:
% actual monthly SSN - 11 years
% half/twice SSN
% velocity:
% actual monthly SSN - 14 years
% not plotted
% density:
% SWPC prediction - 0 years
% SWPC expected SSN ranges
% temperature:
% actual monthly SSN - 14 years
% half/twice SSN
In the beginning of the mission median values of about \SI{87}{\nT}, \SI{340}{\km\per\s}, \SI{214}{\per\cm\cubed} and \SI{503000}{\K} are estimated to be measured at \SI{0.16}{\au}, increasing to about \SI{943}{\nT}, \SI{290}{\km\per\s}, \SI{2951}{\per\cm\cubed} and \SI{1930000}{\K} during the \textbf{first} closest approach at \SI{0.046}{\au}. \textbf{Monthly SSNs---shifted by the time lags specific to the solar wind parameters---are used in the calculation of the solar wind predictions. These SSNs are either actual smoothed values from the SIDC with their reported standard deviations, short-term predictions from the SWPC with their expected ranges or actual smoothed values from the SIDC shifted by 11~years with half/twice their values as uncertainties. The error bands given in both figures, calculated from error propagation, include these SSN ranges and the derived fit parameter errors.}
%B
% 87.3642361846083
%N
% 213.613795298952	first
% 2950.67837375615	nearest
% 2925.63248639046	SSN 0
% 3421.59065204775	SSN 200
%T
% 503136.759398494

\textbf{Finally the estimated solar wind environment can be derived from the function $W'''(x,ssn,r)$. The estimated frequency distributions of the four solar wind parameters at PSP's first and 22nd (first closest) perihelia are plotted in Fig.~\ref{fig:SPP_sw_distributions_b}.}
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/SPP_sw_distributions_b.pdf}}
	\caption{\textbf{Frequency distributions of the four solar wind parameters (same as in Figs.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot} and \ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}) and those estimated with the solar wind model for PSP's first and 22nd (first closest) perihelion. In these figures the frequencies of both extrapolated curves are scaled for visibility to the same height as the \SI{1}{\au} distribution.}}
	\label{fig:SPP_sw_distributions_b}
\end{figure}
\textbf{Note again that the velocity and temperature distributions for the 22nd perihelion are only upper limits and the actual by PSP to be encountered values are expected to be smaller.}

@@

\textbf{For determining solar activity and solar distance dependent relations for the median and average solar wind values, we could have used the simpler approach of combining the radial dependence of averaged Helios data with averaged \SI{1}{\au} OMNI data scaled with the SSN. It is expected that the results of a simpler analysis would have similar distance scaling results, as can be inferred from the exponents in Tables~\ref{tab:mean_median_fit_parameter} and \ref{tab:extrapolation_model_fit_parameters}. However, in our study we are not only interested in averages but rather in bulk distributions, that is, the whole range of values that might occur. For the determination of the frequency distributions the use of the more complex fit model is important, because the distance between median and average values determines the width of the lognormal distributions.}

\textbf{It} is clear that the calculated distribution functions just represent first order estimates of the real solar wind to be encountered by PSP. The solar wind environment to be encountered will depend at times of PSP on the structure of the solar corona and underlying photospheric magnetic field and on the evolution and interaction of individual solar wind streams and superimposed CMEs and shocks. However, the derived results are in good agreement with existing studies about near-Sun solar wind magnetic field strenghts and densities as shown in Sect.~\ref{sec:model_extrapolation_to_psp_orbit}. \textbf{The extrapolation results of the velocity and the temperature differ from the direct measurements seen in existing studies. This suggests that below about \SI{20}{\Rs} PSP may dive into the region where the acceleration and heating of the solar wind is expected to occur} (see Fig.~\ref{fig:sw_extrapolation_ssn_f_plot}). The near-Sun solar wind velocity at PSP perihelion is also expected to be slower than our model estimates, because the \textbf{solar wind is assumed to be accelerated up to the height of the Alfvénic critical surface, which is predicted to lie on average around \SI{17}{\Rs} \citep[e.g.,][]{Sittler1999,Exarhos2000}, scaling with solar activity within a range of between \SI{15}{\Rs} at solar minimum and \SI{30}{\Rs} at solar} maximum \citep{Katsikas2010,Goelzer2014}.	% in average around 17~Rs (e.g., 0.6~Rs Schatten 1969, 17~Rs Sittler1999, 17~Rs Exarhos2000) and scales with solar activity between 15~Rs in minimum and 30~Rs in maximum (19~Rs Katsikas2010, 15–30~Rs Goelzer2014).
